---
title: "Nutrient Concentration of the Devereux Slough 2003-2022"
format: html
editor: visual
---

---
title: "Devereux Slough Nutrient Concentration 2003-2022"
output: html_notebook
---

Agularia data (2003-2011) is available at https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.6.

Nutrient data for this study was taken at different intervals 1-24 meters from Venoco bridge.

Cheadle Center data (2017-2022) is available at https://doi.org/10.25349/D9VG9J

Water years are determined based on an October 1st start date. The average rainfall at UCSB is 16.98 inches https://files.countyofsb.org/pwd/hydrology/historic%20data/rainfall/yearly%20graphs/200graph.pdf.

```{r, echo=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(dplyr)
library(zoo)
setwd("C:/Users/rickard/Documents/test (nutrient r code)")
```

```{r}
# Needed Function 
getYearQuarter <- function(x,
                           firstMonth=10,
                           fy.prefix='WY',
                           quarter.prefix='Q',
                           sep='-',
                           level.range=c(min(x), max(x)) ) {
  if(level.range[1] > min(x) | level.range[2] < max(x)) {
    warning(paste0('The range of x is greater than level.range. Values ',
                   'outside level.range will be returned as NA.'))
  }
  quarterString <- function(d) {
    year <- as.integer(format(d, format='%Y'))
    month <- as.integer(format(d, format='%m'))
    y <- ifelse(firstMonth > 1 & month >= firstMonth, year+1, year)
    q <- cut( (month - firstMonth) %% 12, breaks=c(-Inf,2,5,8,Inf),
              labels=paste0(quarter.prefix, 1:4))
    return(paste0(fy.prefix, substring(y,3,4)))
  }
  vals <- quarterString(x)
  levels <- unique(quarterString(seq(
    as.Date(format(level.range[1], '%Y-%m-01')),
    as.Date(format(level.range[2], '%Y-%m-28')), by='month')))
  return(factor(vals, levels=levels, ordered=TRUE))
} 
```

```{r}
# upload historic rain data for santa barbara
rain<-read_excel("Data/Rain/200dailys.xls",skip=9)
colnames(rain)[4]<-"month"
colnames(rain)[6]<-"Daily_Rain"
rain$Date<-as.Date(paste0(rain$year,"/",rain$month,"/",rain$day),format="%Y/%m/%d")
rain<-rain[!is.na(rain$Date),]
rain$wtr_yr <- getYearQuarter(rain$Date, firstMonth=10)
rain<-rain[rain$`water year`>2002,]
rain<-as.data.frame(rain)
rain_wtryr<-aggregate(Daily_Rain~wtr_yr,rain,FUN=sum)

#Looking at rainfall totals- very different, so going off County webpage instead
rain_wtryr$rainfall <- ifelse(rain_wtryr$Daily_Rain > 18.39,"Wet","dry")
ggplot(rain_wtryr)+geom_col(aes(x=wtr_yr,y=Daily_Rain,fill=rainfall))+theme_bw()
```

Years in blue are above average rainfall, years in red are below average.

```{r, echo=FALSE, warning=FALSE}
#Upload and format old nutrient data from 2004-2011
old_Dev<-read.csv("Data/Nutrients/sbclter_stream_chemistry_allyears_non_registered_stations_20190628.csv")
old_Dev<-old_Dev[grepl("^DV", old_Dev$site_code), ]
old_Dev$Date<-as.Date(old_Dev$timestamp_local,format="%Y-%m-%dT%H:%M:%S")
old_Dev_m<-old_Dev[,c(3:5,13)]
colnames(old_Dev_m)<-c("Ammonia.um","Nitrite.Nitrate.um","Phosphate.um","Date")
old_Dev_m$Environment<-"Golf Course" #naming environment for later merging

#Upload and format new nutrient data from 2017-2022
new_Dev<-read.csv("Data/nutrients/Nutrients_2018-2022.csv",strip.white = TRUE)
new_Dev$Sample.Date<-as.Date(new_Dev$Sample.Date,format="%m/%d/%Y")
new_Dev<-new_Dev[new_Dev$Site=="Venoco"|new_Dev$Site=="Venoco Bridge",]
new_Dev<-new_Dev[,c(2,4:6)]
colnames(new_Dev)[1]<-"Date"
new_Dev$Environment<-"Wetland" #naming environment for later merging
```

```{r}
#Merge old and new data together
Dev_all<-rbind(new_Dev,old_Dev_m)
Dev_all_rain<-merge(Dev_all,rainfull,by="Date",all.x=TRUE)
Dev_all_rain<-Dev_all_rain[Dev_all_rain$Phosphate.um>=0&Dev_all_rain$Nitrite.Nitrate.um>=0&Dev_all_rain$Ammonia.um>=0,]

#Plot data
ggplot(Dev_all_rain)+geom_point(aes(x=twodayrain,y=Ammonia.um, color=Environment))+theme_bw()
ggplot(Dev_all_rain)+geom_boxplot(aes(x=Environment,y=Nitrite.Nitrate.um))+theme_bw()
ggplot(Dev_all_rain)+geom_point(aes(x=twodayrain,y=Nitrite.Nitrate.um, color=Environment))+theme_bw()
ggplot(Dev_all_rain)+geom_boxplot(aes(x=Environment,y=Ammonia.um))+theme_bw()
ggplot(Dev_all_rain)+geom_boxplot(aes(x=Environment,y=Ammonia.um))+theme_bw()+ylim(0,50)
ggplot(Dev_all_rain)+geom_point(aes(x=Daily_Rain,y=Phosphate.um, color=Environment))+theme_bw()
ggplot(Dev_all_rain)+geom_boxplot(aes(x=Environment,y=Phosphate.um))+theme_bw()
Dev_all_rain
```

Beginning summary shows that the median is typically higher after the wetland is established, but there are many more high outlines while it was a golf course. This likely has to do with data collection differences. Golf course had weekly sampling even during the summer, while wetland only sampled during rain storms.

```{r, echo=FALSE, warning=FALSE}
#Summarize data and combine with rain years
Dev_all_rain$wtr_yr <- getYearQuarter(Dev_all_rain$Date, firstMonth=10)
Dev_all_rain<-Dev_all_rain[Dev_all_rain$wtr_yr!="WY03",]
summary<- Dev_all_rain %>% 
  group_by(wtr_yr) %>% 
  summarize(
    avg_P = mean(Phosphate.um),
    med_P = median(Phosphate.um),
    max_P = max(Phosphate.um),
    avg_N = mean(Nitrite.Nitrate.um),
    med_N = median(Nitrite.Nitrate.um),
    max_N = max(Nitrite.Nitrate.um),
    avg_A = mean(Ammonia.um),
    med_A = median(Ammonia.um),
    max_A = max(Ammonia.um)
  )
waterstats<-merge(summary,rain_wtryr,by="wtr_yr")
plot(summary$max_A~summary$wtr_yr)
plot(summary$max_N~summary$wtr_yr)
plot(summary$max_P~summary$wtr_yr)
Dev_all_rain_mg<-Dev_all_rain
Dev_all_rain_mg$P_mg<-(Dev_all_rain_mg$Phosphate.um*30.97)/1000
Dev_all_rain_mg$N_mg<-(Dev_all_rain_mg$Nitrite.Nitrate.um*14.0067)/1000
Dev_all_rain_mg$A_mg<-(Dev_all_rain_mg$Ammonia.um*14.0067)/1000
```

While the median of the wetland is similar, the max values are almost always higher for the golf course.

```{r, echo=FALSE, warning=FALSE}

summary2<- Dev_all_rain_mg %>% 
  group_by(Environment) %>% 
  summarize(
    avg_P = mean(P_mg),
    med_P = median(P_mg),
    max_P = max(P_mg),
    avg_N = mean(N_mg),
    med_N = median(N_mg),
    max_N = max(N_mg),
    avg_A = mean(A_mg),
    med_A = median(A_mg),
    max_A = max(A_mg)
  )
summary2<-as.data.frame(summary2)
#possibly will want to get daily averages.
Dev_all_rain_Daily<-aggregate(cbind(Ammonia.um,Phosphate.um,Nitrite.Nitrate.um)~Date+Daily_Rain+Environment+twodayrain,Dev_all_rain,FUN=mean)

#Plot data- Ammonia
ggplot(Dev_all_rain_Daily)+geom_point(aes(x=Daily_Rain,y=Ammonia.um, color=Environment))+
  theme_bw()

#Plot data- Nitrate
ggplot(Dev_all_rain_Daily)+geom_point(aes(x=Daily_Rain,y=Nitrite.Nitrate.um, color=Environment))+
  theme_bw()

ggplot(Dev_all_rain_Daily)+geom_point(aes(x=Daily_Rain,y=Phosphate.um, color=Environment))+
  theme_bw()+ylim(0,100)


# There are a lot of high nutrient concentrations for the golf course, but also a lot more samples taken now try to seperate by month
Dev_all_rain$month<-format(Dev_all_rain$Date,format="%b")
Dev_all_rain$Year<-format(Dev_all_rain$Date,format="%Y")

ggplot(Dev_all_rain)+geom_boxplot(aes(x=month,y=Phosphate.um, color=Environment))
ggplot(Dev_all_rain)+geom_boxplot(aes(x=month,y=Nitrite.Nitrate.um, color=Environment))
ggplot(Dev_all_rain)+geom_boxplot(aes(x=month,y=Ammonia.um, color=Environment))+ylim(0,100)




## What does old devereux data look like?
old_Dev<-old_Dev[old_Dev$po4_uM>=0&old_Dev$no3_uM>=0&old_Dev$nh4_uM>=0,]
old_Dev$Year<-format(old_Dev$Date,format="%Y")

ggplot(old_Dev)+geom_boxplot(aes(x=site_code,y=po4_uM))
ggplot(old_Dev)+geom_boxplot(aes(x=site_code,y=nh4_uM))
ggplot(old_Dev)+geom_boxplot(aes(x=site_code,y=no3_uM))

ggplot(old_Dev)+geom_boxplot(aes(x=Year,y=po4_uM))
ggplot(old_Dev)+geom_boxplot(aes(x=Year,y=nh4_uM))
ggplot(old_Dev)+geom_boxplot(aes(x=Year,y=no3_uM))

ggplot(Dev_all_rain)+geom_boxplot(aes(x=Year,y=Phosphate.um))
ggplot(Dev_all_rain)+geom_boxplot(aes(x=Year,y=Nitrite.Nitrate.um))
ggplot(Dev_all_rain)+geom_boxplot(aes(x=Year,y=Ammonia.um))
```
